A Component-Based Sensor Network for \\ Environmental Monitoring

### 2009-03-13 docs/icccn2009/paper.tex
Edited section on DSP
git-svn-id: svn://localhost/netbeams/trunk@374 559edde2-a4d4-44e0-96fc-4035e40d7396

In the following we present a practical application of the DSP
presented in the previous section. We describe a deployment of the DSP
using sensor owned and operated by the RTC. Our environmental sensor
network is dubbed NetBEAMS (Networked Bay Environmental Assessment
Monitoring System) and allows access to sensor equipment deployed
throughout the San Francisco Bay Area.  Our goal is to address the
problem of data collection from remote sensors by using the DSP to
automatically interrogate sensors at the remote field locations and
transmit the collected measurements over long distances to a backend
server where the data is made available to potential users in near
real-time. This end-to-end system will be a ready-to-use turnkey
solution that is adaptable to a wide variety of sensor types and
requires no programming expertise to operate. Configuration of the
system will be possible when the sensor is deployed in the field via a
web-based interface, eliminating the need for extra site visits by
technicians between service cycles to change systems functionalities
(e.g. sampling rate). We present the NetBEAMS infrastructure by
discussing its hardware and software components.

@@

Figure \ref{FIG_NETBEAMS} provides an end-to-end overview of the
NetBEAMS architecture. NetBEAMS is built using multiple installations
of the DSP and by providing special purpose components for the various
tasks. One DSP installation is collocated with each sensor and an
additional DSP is running on the backend server. Because of the remote
locations where each individual sensor is deployed in the field and
the sparse nature of the network, the resulting topology is a tree
structure of depth 1.

@@

In our project, we use the Verdex Pro XM4 motherboard with 64MB of
SDRAM, 16MB of Flash, and a 400MHz Marvell PXA270 with XScale
processor.  We also use two expansion boards: the Console-VX which
comes with three RS232 serial ports and the Netpro-VX which features
one 10/100baseT ethernet port.  The Gumstix runs Linux 2.6 with the
BusyBox utilities, and uses the OpenEmbedded cross-compile build
environment to provide a complete Linux environment and a large range
of Linux applications for the ARM architecture.  We use the JamVM
\cite{jamvm01} implementation of Sun Microsystem's virtual machine in
order to run the DSP. The Gumstix communicates with the YSI Sonde via
an RS232 serial port using a null-modem cable since both the Sonde and
the Gumstix are DTE devices. The serial port configuration is:

>>>>

In the following we present a practical application of the DSP
introduced in the previous section. Our environmental sensor network
is dubbed NetBEAMS (Networked Bay Environmental Assessment Monitoring
System) and allows access to sensor equipment deployed throughout the
San Francisco Bay Area. This end-to-end system is a ready-to-use
turnkey solution that is adaptable to a wide variety of sensor types
and requires no programming expertise to operate. Configuration of the
system will be possible when the sensor is deployed in the field via a
web-based interface, eliminating the need for extra site visits by
technicians between service cycles to change systems functionalities
(e.g. sampling rate).  We present the NetBEAMS infrastructure by
discussing its hardware and software components.

@@

Figure \ref{FIG_NETBEAMS} provides an end-to-end overview of the
NetBEAMS architecture. NetBEAMS is built using multiple installations
of the DSP and by providing special purpose components for the various
tasks. One DSP installation is collocated with each sensor and an
additional DSP is running on the backend server. Because of the remote
locations where each individual sensor is deployed in the field and
the sparse density of the network, the resulting topology is a tree
structure of depth 1.

@@

The Gumstix runs Linux 2.6 with the BusyBox utilities, and uses the
OpenEmbedded cross-compile build environment to provide a complete
Linux environment and a large range of Linux applications for the ARM
architecture.  We use the JamVM \cite{jamvm01} implementation of Sun
Microsystem's virtual machine in order to run the DSP. The Gumstix
communicates with the YSI Sonde via an RS232 serial port using a
null-modem cable since both the Sonde and the Gumstix are DTE devices.
The serial port configuration is:
